---
title: "DataVisualization-DesignContest"
author: "RamyaPrakash"
date: "10/17/2019"
output: html_document
#output: 
 #flexdashboard::flex_dashboard:
  #  storyboard: true
   # social: menu
    #source: embed
---

# TidyTuesday:PowerLifting

# DataSet Github Page:https://github.com/rfordatascience/tidytuesday/tree/master/data/2019/2019-10-08 

# Usecase replication Github Page
https://connorrothschild.github.io/tidytuesday/2019-10-08/index



```{r load-packages, message=F, warning=F}
#install.packages("ggplot2")
#install.packages("tidyverse")
#install.packages("cr")
#install.packages("flexdashboard")
library(flexdashboard)
library(ggplot2)
library(tidyverse)
library(tidyr)
#library(cr)

#set_cr_theme(font = "lato")
```

## load the data

```{r load-data}
# df <- readr::read_csv("openpowerlifting-2019-09-20.csv")
# 
# df_clean <- df %>% 
#   janitor::clean_names()
# 
# ipf_lifts <- df_clean %>% 
#   select(name:weight_class_kg, starts_with("best"), place, date, federation, meet_name)  %>% 
#   filter(!is.na(date)) %>% 
#   filter(federation == "IPF")

ipf_lifts <- read_csv("data/ipf_lifts.csv")

```

```{r glimpse}
# run the glimpse() function
glimpse(ipf_lifts)

```

## Clean ipf_lifts, and reshape the three lifts into one column

```{r clean_data-01}
#install.packages("lubridate")

library(lubridate)

ipf_lifts <- ipf_lifts %>% 
  mutate(year = lubridate::year(date))

ipf_lifts_reshape <- ipf_lifts %>% 
  tidyr::pivot_longer(cols = c("best3squat_kg", "best3bench_kg", "best3deadlift_kg"), names_to = "lift") %>% 
  select(name, sex, year, lift, value)

```

## This visualization is concerned with the heaviest lifts from each year

```{r clean_data-02}
ipf_lifts_maxes <- ipf_lifts_reshape %>% 
  group_by(year, sex, lift) %>% 
  top_n(1, value) %>% 
  ungroup %>% 
  distinct(year, lift, value, .keep_all = TRUE)

```

## In order to construct a dumbbell plot, we need both male and female observations in the same row.

```{r clean_data-03}
max_pivot <- ipf_lifts_maxes %>% 
  spread(sex, value)
```

## Now construct a dataframe for each sex and join them:

```{r clean_data-04}
male_lifts <- max_pivot %>% 
  select(-name) %>% 
  filter(!is.na(M)) %>% 
  group_by(year, lift) %>% 
  summarise(male = mean(M))

female_lifts <- max_pivot %>% 
  select(-name) %>% 
  filter(!is.na(`F`)) %>% 
  group_by(year, lift) %>% 
  summarise(female = mean(`F`))

max_lifts <- merge(male_lifts, female_lifts)

max_lifts_final <- max_lifts %>% 
  group_by(year, lift) %>% 
  mutate(diff = male - female)
```

## Construct the visualization.
## First, a static viz

```{r viz-01}
#install.packages("devtools")
#devtools::install_github("clauswilke/ggtext")
#devtools::install_github("connorrothschild/tpltheme")
library(tpltheme)
#install.packages("ggalt")
library(ggtext)
max_lifts_final %>% 
  filter(year == 2019) %>% 
  ggplot() + 
  ggalt::geom_dumbbell(aes(y = lift,
                    x = female, xend = male),
                colour = "grey", size = 5,
                colour_x = "#D6604C", colour_xend = "#395B74") +
  labs(y = element_blank(),
       x = "Top Lift Recorded (kg)",
       title =  "How <span style='color:#D6604C'>Women</span> and <span style='color:#395B74'>Men</span> Differ in Top Lifts",
       subtitle = "In 2019") +
  theme(plot.title = element_markdown(lineheight = 1.1, size = 20),
        plot.subtitle = element_text(size = 15)) +
  scale_y_discrete(labels = c("Bench", "Deadlift", "Squat")) +
  drop_axis(axis = "y") +
  geom_text(aes(x = female, y = lift, label = paste(female, "kg")),
            color = "#D6604C", size = 4, vjust = -2) +
  geom_text(aes(x = male, y = lift, label = paste(male, "kg")),
            color = "#395B74", size = 4, vjust = -2) +
  geom_rect(aes(xmin=430, xmax=470, ymin=-Inf, ymax=Inf), fill="grey80") +
  geom_text(aes(label=diff, y=lift, x=450), fontface="bold", size=4) +
  geom_text(aes(x=450, y=3, label="Difference"),
                     color="grey20", size=4, vjust=-3, fontface="bold")
```

#Column {data-width=650}
#-------------------------------------
    
## Add animation

```{r viz-animation-01}
#install.packages('gganimate')
#install.packages("gifski")
library(gganimate)
library(gifski)
animation <- max_lifts_final %>% 
  ggplot() + 
  ggalt::geom_dumbbell(aes(y = lift,
                    x = female, xend = male),
                colour = "grey", size = 5,
                colour_x = "#D6604C", colour_xend = "#395B74") +
  labs(y = element_blank(),
       x = "Top Lift Recorded (kg)",
       title =  "How <span style='color:#D6604C'>Women</span> and <span style='color:#395B74'>Men</span> Differ in Top Lifts",
       subtitle='\nThis plot depicts the difference between the heaviest lifts for each sex at International Powerlifting Federation\nevents over time. \n \n{closest_state}') +
  theme(plot.title = element_markdown(lineheight = 1.1, size = 25, margin=margin(0,0,0,0)),
        plot.subtitle = element_text(size = 15, margin=margin(8,0,-30,0))) +
  scale_y_discrete(labels = c("Bench", "Deadlift", "Squat")) +
  drop_axis(axis = "y") +
  geom_text(aes(x = female, y = lift, label = paste(female, "kg")),
            color = "#D6604C", size = 4, vjust = -2) +
  geom_text(aes(x = male, y = lift, label = paste(male, "kg")),
            color = "#395B74", size = 4, vjust = -2) +
  transition_states(year, transition_length = 4, state_length = 1) +
  ease_aes('cubic-in-out')

a_gif <- animate(animation, 
                 fps = 10, 
                 duration = 25,
        width = 800, height = 400, 
        renderer = gifski_renderer("./heavy_lifts_each_sex.gif"))

a_gif
```

#Column {data-width=350}
-------------------------------------
## Add animation

```{r viz-animation-02}
animation2 <- max_lifts_final %>% 
  ungroup %>% 
  mutate(lift = case_when(lift == "best3bench_kg" ~ "Bench",
                          lift == "best3squat_kg" ~ "Squat",
                          lift == "best3deadlift_kg" ~ "Deadlift")) %>% 
  ggplot(aes(year, diff, group = lift, color = lift)) + 
  geom_line(show.legend = FALSE) + 
  geom_segment(aes(xend = 2019.1, yend = diff), linetype = 2, colour = 'grey', show.legend = FALSE) + 
  geom_point(size = 2, show.legend = FALSE) + 
  geom_text(aes(x = 2019.1, label = lift, color = "#000000"), hjust = 0, show.legend = FALSE) + 
  drop_axis(axis = "y") +
  transition_reveal(year) +
  coord_cartesian(clip = 'off') +
  theme(plot.title = element_text(size = 20)) +
  labs(title = 'Difference over time',
       y = 'Difference (kg)',
       x = element_blank()) + 
  theme(plot.margin = margin(5.5, 40, 5.5, 5.5))

b_gif <- animate(animation2, 
                 fps = 10, 
                 duration = 25,
        width = 800, height = 200, 
        renderer = gifski_renderer("./difference_over_time.gif"))

b_gif
```

## Combine animation: engine crashing-code commented below TBD

```{r viz-animation-03}
#install.packages("magick")
library(magick)
a_mgif <- image_read(a_gif)
b_mgif <- image_read(b_gif)

new_gif <- image_append(c(a_mgif[1], b_mgif[1]), stack = TRUE)
for(i in 2:250){
  combined <- image_append(c(a_mgif[i], b_mgif[i]), stack = TRUE)
  new_gif <- c(new_gif, combined)
}

new_gif
```

## RamyaP visuals: 

```{r my-viz}
#install.packages("ggridges")
library(gganimate)
library(gifski)
library(ggridges)
ipf_lifts_year <- ipf_lifts %>% 
                  mutate(year = format(date, "%Y")) %>%
                  filter(year %in% c(2009:2019))

ipf_lifts_decade<- ggplot(data=ipf_lifts_year, mapping = aes(x=best3squat_kg, y=year, fill=sex)) + 
geom_density_ridges() + 
scale_fill_manual(values = colorspace::darken(c("#E69F00", "#56B4E9", "#009E73", "#F0E442", "#0072B2", "#999999")), labels = c("female", "male")) +
scale_color_manual(values = colorspace::darken(c("#E69F00", "#56B4E9", "#009E73", "#F0E442", "#0072B2", "#999999"), 0.3))+
labs(x = 	"Weight (kg)",y = "Year",title = "Squat") +
scale_x_continuous(limits = c(10,500)) + theme_ridges()+ transition_manual(year)

my_gif <- animate(ipf_lifts_decade, 
                 fps = 5, 
                 duration = 10,
        width = 800, height = 200, 
        renderer = gifski_renderer("./ipf_lifts_decade.gif"))

my_gif

```